\documentclass[main]{subfiles}

\begin{document}
\chapter*{Preface to New Typesetting}
\addcontentsline{toc}{chapter}{Preface to New Typesetting}

This book is a re-LaTeX'ed version of the book ``Notes on Differential Geometry'' by Hicks. We're the \TeX
{}romancers, a group of mostly math enthusiasts wanting to create more readable content for the math world. The typesetting credits go to: Aareyan Manzoor, George Coote, RokettoJanpu, Carl Sun, Andrew Lin, Nathaniel Alloway, Prakhar Agarwal, Shuayb Mohammed, Bastián Núñez, Arpit Mittal. %add your name here

Here is a link to a dyslexia friendly version of the book: \url{https://aareyanmanzoor.github.io/assets/hicks-dyslexic.pdf}

Some notations in the original book are rather ancient or just not as popular, so we decided to change these to their modern counterpart. For example, set builders are written using $[~]$ in the original book; we decided to use $\{\}$. We also used $\tangentspace{M}{m}$, $\tangentbundle{M}$ for the tangent space at $m$, the tangent bundle of $M$, rather than $M_m$ and $T(M)$. We further used $\formbundle{M}{p}$ to denote the space of $p$-forms, and $\tensorbundle{}{p}{q}V$ to denote mixed tensors over some vector space $V$. We changed $L_XY$ to $\liederiv{X}Y$ for the Lie derivative. However, we decided to keep the name \emph{connexion} instead of the modern name \emph{connection} as it has some charm.

The original book had a lot of inline equations to its detriment, and we decided to display mode some of these. A lot of things listed were made into an itemized list also. Some sentences were changed for better readability. A lot of the problems were split up into parts labeled by roman numbers.

We added citations and references with hyperlinks. References to e.g. theorems in the book are in blue, while citations to the bibliography is in red. The bibliography also has URLs now, for easy access. Some of the books in the bibliography  had newer editions, so we went with those.

All the diagrams have been redrawn, special thanks to our artists: Yohan Wittgenstein, Bastián Núñez, John Cerkan. %add your names here if you want, yohan derivada and jcerkan. 






\end{document}